### 2024-05-21 thesis/abstract.tex
May 21st 2024

\begin{abstract}
    The theory of kernel methods will be applied to the problem of probabilistic forecasting taking as input some data.
    Conceptually, the proposed analysis could be applied to any kind of data. In this study we considered the electricity market, because of the interesting implications  on risk management tasks.

  %This example thesis briefly shows the main features of our thesis
  %style, and how to use it for your purposes.
\end{abstract}

>>>>

\begin{abstract}
  % introduction
  This thesis is concerned with electricity forecasting. 
  Specifically, we concentrated on the probabilistic framework. This choice was motivated by its importance in systems planning and operations as a consequence of the inception of competitive power markets, smart grids and renewable integration requirements.
  % methods
  In doing so, we focused on the family of kernel methods. In addition, we also compared them against several statistical and machine learning techniques.
  % results
  Our results showed the goodness of kernel methods in the field of electricity forecasting both in terms of point and probabilistic forecasting. In particular, in the probabilistic context, our experiments showed the validity of kernel quantile regression equipped with the absolute Laplacian kernel.
  % discussion
  These findings indicate kernel methods are well suited to the characteristic of electricity.
  Anyone interested in forecasting energy should consider them when faced with the choice of the model.
  They can be employed stand alone or combined with other valid methods into ensembles.

  %This example thesis briefly shows the main features of our thesis
  %style, and how to use it for your purposes.
\end{abstract}
  % The theory of kernel methods will be applied to the problem of point and probabilistic forecasting for the energy sector.
  %   Such choice is motivated by its interesting implications.

### 2024-05-21 thesis/input_files/analysis.tex
May 21st 2024

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover.
% figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

\subsection{Long short term memory}
We used the torch library in order to code our LSTM regressor.
Various hyperparameters combinations were tried during hyperparameter tuning. The final hyperparameters we set on follows.
\begin{itemize}
    \item hidden\_size= 64
    \item num\_layers= 2
    \item output\_size= 1
    \item num\_epochs= 30
    \item learning\_rate= 0.001
    \item batch\_size= 32
    \item window\_size= 24
\end{itemize}
LSTM forecast is reported in table \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/lstm_price.png}
    \caption{Long short term memory prediction}
    \label{fig:lstm_price}
\end{figure}

>>>>

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependent variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}.
% , moreover figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet model\_2 prediction}
    \label{fig:prophet_price2}
\end{figure}

@@

\subsection{Long short term memory}
We used the torch library in order to code our LSTM regressor.
Various hyperparameters combinations were tried during the tuning phase. The final architecture we set on follows
\begin{itemize}
    \item hidden\_size= 64
    \item num\_layers= 2
    \item output\_size= 1
    \item num\_epochs= 30
    \item learning\_rate= 0.001
    \item batch\_size= 32
    \item window\_size= 24
\end{itemize}
LSTM forecast is reported in figure \ref{fig:lstm_price}, it achieves a RMSE of 9.1782
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/lstm_price.png}
    \caption{Long short term memory prediction}
    \label{fig:lstm_price}
\end{figure}

### 2024-05-21 thesis/input_files/implementation.tex
May 21st 2024

This section is inteneded to explain and aid for reproducibilty studies. Hereafter the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

Section documenting code

- Explain how methods' implementation has been
adapted to my specific setting.

- Explain in detail how to my src code has been implemented
its rationale and how to use it.

- As I explain code scripts go over the test, to 
explain better my ideas.

- Indicate also hyperparameters maybe in each subsection

@@

\subsection{Tbats}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

@@

\subsection{LSTM}
The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.

@@

\subsection{Kernel support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying rbf as the kernel parameter.

>>>>

This section is inteneded to explain and aid for reproducibilty studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

% Section documenting code

% - Explain how methods' implementation has been
% adapted to my specific setting.

% - Explain in detail how to my src code has been implemented
% its rationale and how to use it.

% - As I explain code scripts go over the test, to 
% explain better my ideas.

% - Indicate also hyperparameters maybe in each subsection

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

@@

\subsection{Long short term memory}
The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.

@@

\subsection{Kernel support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used.
% by specifying rbf as the kernel parameter.
